### 2024-11-26 paper/paper.qmd
Title and author update - paper

---
title: "Analyzing Sakura Flowering Trends Under Climate Change"
subtitle: "Predicting Bloom Dates Using Historical Records and Temperature Data"
author: 
  - Mariko Lee
thanks: "Code and data are available at: [https://github.com/leemarik/sakura-flowering-trends.git](https://github.com/leemarik/sakura-flowering-trends.git)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```
```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace setup ####
library(broom.mixed)
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(here)
library(kableExtra)
library(arrow)
library(rstanarm)

#### Load necessary data and models ####
clean_sakura_data <- read_parquet("../data/02-analysis_data/clean_sakura_data.parquet")
lm_mode <- readRDS("../models/lm_model.rds")
gam_model <- readRDS("../models/gam_model.rds")

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Talk more about it.

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

\newpage

# Model {#sec-model}
To determine the impact of temperature on sakura flowering dates over centuries and develop predictive models for future trends under different climate scenarios, I used a linear regression model (LM) as a baseline and a generalized additive model (GAM) for capturing non-linear relationships in the data.

## Linear Regression Model 
The linear regression model I will be using is:

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

## Generalized Additive Model (GAM)

Where:

### Model justification

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# References

### 2024-11-28 paper/paper.qmd
Adding Abstract and Introduction

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study uses historical and modern data to analyze the impact of temperature and long-term trends on sakura flowering in Japan.
+Linear regression highlights a significant shift toward earlier flowering dates, while generalized additive models (GAMs) capture non-linear relationships.
+Model comparison via AIC favors the GAM, showing its effectiveness in handling complex patterns.
+Predictions under future climate scenarios (+1°C, +2°C, +3°C) indicate flowering dates may advance by 5–15 days by 2070.
+These results underscore the sensitivity of sakura flowering to climate change and its broader ecological and cultural implications."
 format: pdf
@@
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
-library(palmerpenguins)
-```
 ```{r}
@@
 # Introduction
+In Japan, cherry blossoms, or “sakura,” have been a significant cultural and historical symbol, representing the transient beauty of life.
+The practice of “hanami,” or flower viewing, dates back centuries.
+During the Heian period (794-1185), aristocrats would gather to appreciate the blossoms’ fleeting beauty while composing poetry (@japanSakuraHistory).
+Hanami has developed into a deeply ingrained cultural tradition that transcends social classes, with its economic significance growing in modern times.
+Recent research by Katsuhiro Miyamoto indicates that the economic impact of hanami in Japan is expected to double this year, highlighting its role as a cultural event and an economic activity (@kaneko2024hanami).
 
-Overview paragraph
+However, a concerning trend has surfaced in recent decades: sakura blossoming earlier each year, attributed mainly to rising global temperatures.
+According to The Economist, the earlier blooming of cherry blossoms is closely tied to climate change, raising concerns about the long-term implications for Japan’s cultural heritage and biodiversity (@EconomistCherryBlossoms).
+This shift in flowering trends serves as a reminder of humanity’s impact on the environment and a pressing scientific challenge to understand the intricate relationships between climate variables and flowering dates.
 
-Estimand paragraph
+The estimand of this analysis is the relationship between average March temperatures, year, and the Sakura flowering dates.
+By analyzing centuries of historical and modern sakura data, my goal is to determine how changes in temperature and long-term climate trends have impacted the timing of sakura blooming.
+Additionally, I aim to forecast future flowering dates under various climate scenarios, providing insights into the potential effects of climate change on this culturally significant phenomenon.
 
-Results paragraph
+Despite the historical documentation of sakura flowering trends, there is a critical gap in understanding how these changes will evolve in the future under various climate scenarios.
+This analysis addresses this gap by using historical and modern data on sakura flowering dates and temperatures to determine the impact of climate change on flowering dates.
+I used statistical modeling techniques of linear regression and generalized additive models (GAMs) to explore the relationship between temperature, year, and flowering dates.
+These models provide insights into past trends and enable predictions for future scenarios under different temperature increases.
 
-Why it matters paragraph
+The findings indicate that rising temperatures strongly correlate with earlier blooming dates, which have accelerated over the last century.
+By modeling these patterns, the analysis offers a framework for predicting future sakura blooming dates and assessing the broader implications of climate change on cultural and ecological systems.
+This research contributes to scientific understanding and highlights the urgency of climate mitigation efforts to preserve cherished cultural practices like hanami.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The structure of this paper is as follows: @sec-data reviews the data sources and preprocessing methods used in this analysis.
+@sec-model details the modeling approaches, specifically Linear Regression and Generalized Additive Models.
+@sec-results presents the results, followed by a discussion in @sec-discussion on the implications, limitations, and future directions.
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
 ## Generalized Additive Model (GAM)
@@
 
-# Results
+# Results {#sec-results}
 
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+# Discussion {#sec-discussion}
 

### 2024-11-29 scripts/05-exploratory_data_analysis.R
EDA update

@@ -10,73 +10,120 @@
 #### Workspace setup ####
 library(tidyverse)  
 library(arrow)
-library(ggplot2) 
+library(patchwork)
+library(ggplot2)
 
 #### Read data ####
 clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
 
 #### Summary Statistics ####
-# General overview of the dataset
-glimpse(clean_sakura_data)
-
-# Summary of numerical variables
-summary(clean_sakura_data)
-
-# Count of observations by source
-clean_sakura_data %>%
-  count(source) %>%
-  mutate(percentage = n / sum(n) * 100)
-
-# Check for missing values
-clean_sakura_data %>%
-  summarise(across(everything(), ~ sum(is.na(.)))) %>%
-  pivot_longer(cols = everything(), names_to = "variable", values_to = "missing_count")
+# Summary of the key variables
+summary_stats <- clean_sakura_data %>%
+  summarize(
+    min_year = min(year),
+    max_year = max(year),
+    avg_temp_mean = mean(avg_temperature, na.rm = TRUE),
+    avg_temp_sd = sd(avg_temperature, na.rm = TRUE),
+    day_of_year_mean = mean(day_of_year, na.rm = TRUE),
+    day_of_year_sd = sd(day_of_year, na.rm = TRUE)
+  )
+print(summary_stats)
 
 #### Visualizations ####
 
-# 1. Trend of Flowering Dates Over Time
-ggplot(clean_sakura_data, aes(x = year, y = day_of_year, color = source)) +
-  geom_point(alpha = 0.6) +
-  geom_smooth(method = "loess", se = FALSE) +
-  labs(title = "Trend of Sakura Flowering Dates Over Time",
-       x = "Year", y = "Day of Year (Flowering Date)") +
+# 1. Distribution of Flowering Dates (day_of_year)
+plot_flowering_distribution <- ggplot(clean_sakura_data, aes(x = day_of_year)) +
+  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
+  labs(
+    title = "Distribution of Sakura Flowering Dates",
+    x = "Day of Year (Flowering Date)",
+    y = "Frequency"
+  ) +
   theme_minimal()
 
-# 2. Relationship Between Temperature and Flowering Dates
-ggplot(clean_sakura_data, aes(x = avg_temperature, y = day_of_year, color = source)) +
-  geom_point(alpha = 0.6) +
[…]

### 2024-12-02 scripts/07-train_data.R
add training and testing data file and code

@@ -1,0 +1,30 @@
+#### Preamble ####
+# Purpose: testing and training data
+# Author: Mariko Lee
+# Date: 19 November 2024
+# Contact: mariko.lee@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+
+#### Workspace setup ####
+library(dplyr)
+library(rsample)
+library(arrow)
+
+# Read the cleaned sakura dataset
+clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
+
+# Set seed for reproducibility
+set.seed(123)  
+
+# Split the data into 80% training and 20% testing
+split <- initial_split(clean_sakura_data, prop = 0.8)
+
+# Extract training and testing datasets
+train_data <- training(split)
+test_data <- testing(split)
+
+# Save training and testing datasets as Parquet files
+write_parquet(train_data, "data/03-model_data/train_data.parquet")
+write_parquet(test_data, "data/03-model_data/test_data.parquet")

### 2024-12-02 paper/paper.qmd
paper update - half

@@
 library(kableExtra)
+library(patchwork)
 library(arrow)
 library(rstanarm)
+library(Metrics)
+library(mgcv)
 
@@
 # Data {#sec-data}
+## Overview
+This study uses the statistical programming language `R` [@citeR] to analyze sakura flowering patterns across Japan over the centuries, integrating three datasets obtained from Alex Cookson’s repository [@AlexSakura20].
+These datasets include historical Sakura flowering records (812-2015) from Kyoto, modern Sakura flowering data (1953-2019) from across Japan, and temperature data sourced from the Japan Meteorological Agency [@JapanMeteorAgency].
 
-## Overview
+The historical dataset includes reconstructed flowering dates based on Kyoto hanami records and paleoclimatological studies, with estimated March temperature.
+The modern dataset provides nationwide observations of flowering and full bloom dates alongside geographic and temperature data.
+Temperature data for March was sourced directly from Japan Meteorological Agency’s climate statistics.
 
-Overview text
+The three datasets were cleaned, standardized, and merged into a unified structure.
+Missing values were removed, flowering dates were converted into day-of-year (DOY) format, and March temperatures were imputed using reconstructed values where necessary.
+New variables such as flowering range (early, mid, late) and decade grouping were created for this analysis.
+
+Key R packages used include `tidyverse` [@citeTidyverse], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `readr` [@citeReadr], `arrow` [@citeArrow].
 
 ## Measurement
+The dataset used in this study was sourced from Alex Cookson's repository, which integrates historical and modern sakura flowering data with temperature observations.
+The historical data originate from Kyoto, with records spanning 812 to 2015 CE, compiled from diary entries, literary works, and paleoclimatology studies.
+The modern dataset, spanning 1953 to 2019, was obtained from the Japan Meteorological Agency and includes data collected across Japan, such as flowering and full bloom dates, geographic locations, and observed temperatures.
+The integrated dataset consists of 3,526 observations after cleaning and merging.
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The cleaned dataset comprises variables such as:
+- `Year`: The year of observation.
+- `Flowering Date`: The calendar date of initial sakura flowering, converted to day-of-year format.
+- `Average Temperature`: March temperatures either observed or reconstructed (historical).
+- `Day of Year`: Numerical representation of the flowering date within the year. 
+- `Decade`: Grouped decade for temporal trends.
+- `Flowering Range`: Categorical classification of flowering timing (Early, Mid, Late).
+- `Source`: Indicates the origin of the data (modern or historical).
 
-## Outcome variables
+Historical temperature data for Kyoto were reconstructed using paleoclimatology techniques, providing critical context for trends in sakura flowering dates.
+While this reconstruction supports long-term trend analysis, it introduces potential biases due to variations in methodology and accuracy over time.
+Similarly, modern data include direct observations of March temperatures and flowering times recorded at multiple locations, offering broader geographic coverage but differing from the localized Kyoto records in the historical dataset.
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+To ensure comparability, the flowering and temperature data were transformed into consistent numeric formats, allowing analysis of trends over time.
+Missing or incomplete entries were removed, and observed temperatures were prioritized over reconstructed values when available.
+The final dataset was prepared for analysis with derived variables to capture temporal and categorical trends in flowering dates.
 
-Talk more about it.
+## Limitations
+Although the dataset allows long-term sakura flowering trend analysis, several limitations must be noted:
 
-Talk way more about it.
+- Reconstructed Data: Historical temperature data for Kyoto rely on paleoclimatological reconstructions, which may introduce biases or inaccuracies due to methodological differences.
+- Geographic Variability: The historical data is Kyoto-specific, while modern data represents nationwide observations. Differences in regional climate conditions may impact direct comparisons.
+- Temporal Gaps: The historical dataset spans over a millennium but includes periods with sparse or no data, particularly for earlier centuries.
+- Consistency of Definitions: Variations in how flowering dates were recorded historically versus modern observation standards may affect the accuracy of trend analysis.
 
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+Despite these limitations, the dataset analyzes the relationship between climate change and sakura flowering patterns across Japan.
 
@@
 # Model {#sec-model}
-To determine the impact of temperature on sakura flowering dates over centuries and develop predictive models for future trends under different climate scenarios, I used a linear regression model (LM) as a baseline and a generalized additive model (GAM) for capturing non-linear relationships in the data.
+My modeling process aimed to predict the sakura flowering dates (DOY) under varying climate conditions and identify critical factors influencing flowering patterns over time.
+To achieve this, I used both linear regression and generalized additive models (GAMs) to capture relationships between sakura flowering dates, temperature, and time.
+These models were implemented in `R` using the `lm` [@citeLM] and `mgcv` [@citeGAM] packages.
+
+Given the historical and modern data, capturing variability and trends needed the consideration of linear and non-linear relationships.
+While the linear regression model offered a straightforward approach, the GAM allowed flexibility in accounting for complex, non-linear trends.
+Both models were evaluated to ensure a balance between simplicity and accuracy.
 
 ## Linear Regression Model 
-The linear regression model I will be using is:
+The first step in my analysis was building a linear regression model to establish baseline relationships between flowering dates, average March temperatures, and year. This model was expressed as:
+
+$$ \text{day\_of\_year}_i = \beta_0 + \beta_1 \cdot \text{avg\_temperature}_i + \beta_2 \cdot \text{year}_i + \epsilon_i $$
+Where:
+- $\text{day\_of\_year}_i$: Day of Year for the $i^{th}$ observation (flowering date).
+- $\text{avg\_temperature}_i$: Average March temperature.
+- $\text{year}_i$: Year of observation.
+- $\beta_0, \beta_1, \beta_2$: Coefficients for intercept, temperature, and year, respectively.
+- $\epsilon_i$: Random error term.
+
+This model provided a starting point to explore how flowering dates shifted over time.
+The coefficients revealed the effect of each variable, while residual diagnostics were used to evaluate the model's assumptions, such as linearity and homoscedasticity.
+Despite its simplicity, the linear model struggled to capture the non-linear trends evident in the dataset.
 
 ## Generalized Additive Model (GAM)
+To address the limitations of linear regression, I employed a generalized additive model (GAM) to capture non-linear relationships between the predictors and flowering dates.
+The GAM was implemented using the `mgcv` package in R with the following expression:
 
-Where:
+$$ \text{day\_of\_year}_i = \beta_0 + s(\text{avg\_temperature}_i) + s(\text{year}_i) + \epsilon_i $$
 
-### Model justification
+Where $s(\text{avg\_temperature}_i)$ and $s(\text{year}_i)$ are smooth functions capturing non-linear effects of temperature and year.
+These smooth terms were estimated using splines, allowing the model to adjust flexibly to changes in the dataset over time.
+
+This approach offered several advantages, including flexibility, where smooth terms allowed the model to adapt to non-linear relationships, robustness in terms of the REML method, ensuring reliability in estimating smoothness in parameters, and interpretability by visualizing smooth terms, where I was able to identify how temperature and year independently influenced flowering dates.
+
+## Model justification
+The goal of this analysis was to model and predict the sakura flowering dates and explore how they are influenced by temperature and long-term temporal trends.
+To address this, I combined linear regression and GAM model to reflect the goal of balancing interpretability while capturing non-linear trends.
+
+The linear regression model was chosen as the baseline to provide a straightforward interpretation of the relationships between sakura flowering dates, average temperature, and year.
+By incorporating temperature and year as predictors, the model captured the direct effect of warming trends and temporal changes on flowering dates.
+However, its simplicity limited its ability to capture non-linearities of the data.
+
+The generalized additive model was then used to address the limitations of the linear model, precisely its inability to model non-linear trends.
+The smooth terms in GAMs offered the flexible modeling of the relationship between predictors and the response variable.
+GAMs are particularly well-suited for ecological datasets, where non-linear relationships are common due to factors such as thresholds and seasonal variations.
+The inclusion of smooth terms for temperature $s(\text{Temperature})$ and year $s(\text{year})$ model to adapt to varying relationships across the dataset without overfitting.
+
+By comparing the two models, the GAM was a better choice for capturing the complex relationship of sakura flowering trends, as evidenced by its superior fit and lower AIC.
+
+## Model Evaluation and Validation
+To ensure the reliability of the models, I performed the following evaluation and validation steps:
+
+**Model Performance Metrics:** 
+For the linear regression model, the adjusted `R^2` was 0.0136, indicating a weak fit to the data.
+The residual standard error was 28.18 days, highlighting high variability in flowering dates not explained by the model.
+The GAM improved the adjusted `R^2` to 0.0411, explaining 4.36% of the deviance.
+While small, this improvement shows the GAM’s ability to capture non-linear trends lacking in the linear model.
+
+**Diagnostics:**
+Residual plots showed a non-random pattern, suggesting the presence of a non-linear relationship and the need for a more flexible model in the linear regression model.
+For the GAM, the residual checks confirmed model convergence, with all smooth terms for `avg_temperature` and `year` showing significant contributions (p<0.001).
+The basis dimension tests indicated no under-smoothing, making sure that the model captured the complexity of the data.
+
+**Model Comparison:**
+The Aikake Information Criterion (AIC) favored the GAM (AIC = 33466.71) over the linear model (AIC = 33555.97), confirming the GAM’s superior fit.
+The GAM’s flexibility allows non-linear trend modeling, providing a more accurate understanding of the relationships in the data.
+
+**Validation:**
+The dataset was split into training and testing subsets to evaluate predictive performance.
+The dataset was split into a training set (80%) and a testing set (20%).
+Both models were fitted to the training data and evaluated on the testing data using Root Mean Squared Error (RMSE) and Mean Absolute Error (MAE) metrics.
+
+The GAM slightly outperformed the linear regression model, as indicated by lower RMSE and MAE values for both training and testing datasets.
+This aligns with its ability to capture non-linear relationships between predictors and sakura flowering dates.
+Both models presented similar errors on training and testing datasets, suggesting good generalization with no evidence of overfitting.
+The GAM's better performance, especially on testing data, validates its flexibility in modeling complex relationships.
+
+## Future Predictions
+To explore the impact of climate change on sakura flowering dates, the GAM was extended to simulate hypothetical future scenarios.
+These scenarios predict the timing of sakura flowering under varying levels of temperature increases over the next 50 years.
+This approach provides critical insights into the sensitivity of sakura phenology to climate change.
+
+Future predictions were based on three hypothetical warming scenarios:
+- +1°C: A moderate temperature increase.
+- +2°C: A significant increase in average temperature.
+- +3°C: A severe climate change scenario.
+
+The scenarios were created by extrapolating 50 years ahead of the latest recorded year in the dataset and incrementing the average March temperature by 1°C, 2°C, and 3°C.
+The GAM was then used to predict flowering dates for each scenario.
 
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-## Diagnostics
-
 \newpage

### 2024-12-02 paper/paper.qmd
check render ability

@@
 
+```{r}
+#| label: fig-flowering-distribution
+#| fig-cap: Distribution of Sakura Flowering Dates
+#| warning: false 
+#| message: false
+#| fig-align: center
+
+plot_flowering_distribution <- ggplot(clean_sakura_data, aes(x = day_of_year)) +
+  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
+  labs(
+    title = "Distribution of Sakura Flowering Dates",
+    x = "Day of Year (Flowering Date)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+print(plot_flowering_distribution)
+```
+
 # Discussion {#sec-discussion}

### 2024-12-02 paper/paper.qmd
Results section draft update

@@
 
+## Distribution of Flowering Dates and Temperatures
+
+To understand the overall distribution of sakura flowering dates, I’ve plotted the frequency of flowering days across the dataset.
+@fig-flowering-distribution shows that the majority of flowering events occur around the 105th day of the year, corresponding to April 15th.
+This distribution reflects the sensitivity of sakura to climatic conditions in spring.
+
+@fig-temperature-distribution represents the distribution of average March temperature showing a concentration around 5–10°C.
+This temperature range aligns closely with the optimal conditions for sakura blooming, strengthening the relationship between temperature and phenological events.
+
 ```{r}
 #| label: fig-flowering-distribution
-#| fig-cap: Distribution of Sakura Flowering Dates
+#| fig-cap: "Distribution of Sakura Flowering Dates"
 #| warning: false 
@@
 
-print(plot_flowering_distribution)
+plot_flowering_distribution
+```
+```{r}
+#| label: fig-temperature-distribution
+#| fig-cap: "Distribution of Average Temperatures in March"
+#| warning: false 
+#| message: false
+#| fig-align: center
+
+clean_sakura_data_filtered <- clean_sakura_data %>%
+  filter(avg_temperature >= 2 & avg_temperature <= 10)
+
+ggplot(clean_sakura_data_filtered, aes(x = avg_temperature)) +
+  geom_histogram(binwidth = 1, fill = "coral", color = "black") +
+  labs(
+    title = "Distribution of Average Temperatures",
+    x = "Average Temperature (°C)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+```
+## Historical Trends in Flowering Dates and Temperatures
+
+Using @fig-flowering-trend, we see a gradual shift in flowering dates over time.
+Earlier flowering events have become more common, particularly in modern years.
+This trend, captured by the loess regression, suggests an accelerating response to environmental changes.
+Similarly, @fig-temperature-trend reveals a steady increase in average March temperatures over the centuries, with notable warming in recent decades.
+
+```{r}
+#| label: fig-flowering-trend
+#| fig-cap: "Trend of Sakura Flowering Dates Over Time"
+#| warning: false 
+#| message: false
+#| fig-align: center
+
+plot_flowering_trend <- ggplot(clean_sakura_data, aes(x = year, y = day_of_year)) +
+  geom_point(alpha = 0.5, color = "darkgreen") +
+  geom_smooth(method = "loess", color = "blue", se = TRUE) +
+  labs(
+    title = "Trend of Sakura Flowering Dates Over Time",
+    x = "Year",
+    y = "Day of Year (Flowering Date)"
+  ) +
+  theme_minimal()
+
+plot_flowering_trend
+```
+
+```{r}
+#| label: fig-temperature-trend
+#| fig-cap: "Trend of Average Temperatures Over Time"
+#| warning: false 
+#| message: false
+#| fig-align: center
+#| 
+plot_temperature_trend <- ggplot(clean_sakura_data, aes(x = year, y = avg_temperature)) +
+  geom_point(alpha = 0.5, color = "darkorange") +
+  geom_smooth(method = "loess", color = "red", se = TRUE) +
+  labs(
+    title = "Trend of Average Temperatures Over Time",
+    x = "Year",
+    y = "Average Temperature (°C)"
+  ) +
+  theme_minimal()
+plot_temperature_trend
+
+```
+
+## Relationship Between Temperature and Flowering Dates
+
+```{r}
+#| label: fig-temp-vs-flowering
+#| fig-cap: "Relationship Between Temperature and Flowering Date"
+#| warning: false 
+#| message: false
+#| fig-align: center
+
+plot_temp_vs_flowering <- ggplot(clean_sakura_data, aes(x = avg_temperature, y = day_of_year)) +
+  geom_point(alpha = 0.5, color = "purple") +
+  geom_smooth(method = "lm", color = "black", se = TRUE) +
+  labs(
+    title = "Relationship Between Temperature and Flowering Date",
+    x = "Average Temperature (°C)",
+    y = "Day of Year (Flowering Date)"
+  ) +
+  theme_minimal()
+
+plot_temp_vs_flowering
+
+```
+
+## Historical vs. Modern Data Comparison
+
+```{r}
+#| label: fig-flowering-source
+#| fig-cap: "Comparison of Flowering Dates by Data Source"
+#| warning: false 
+#| message: false
+#| fig-align: center
+
+plot_flowering_source <- ggplot(clean_sakura_data, aes(x = source, y = day_of_year, fill = source)) +
+  geom_boxplot() +
+  labs(
+    title = "Comparison of Flowering Dates by Data Source",
+    x = "Source",
+    y = "Day of Year (Flowering Date)"
+  ) +
+  theme_minimal() +
+  scale_fill_manual(values = c("historical" = "lightblue", "modern" = "lightgreen"))
+
+plot_flowering_source
+
+```
+
+## Flowering Ranges Across Decades
+
+```{r}
+#| label: fig-flowering-decade
+#| fig-cap: "Flowering Range by Decade"
+#| warning: false 
+#| message: false
+#| fig-align: center
+
+plot_flowering_decade <- clean_sakura_data %>%
+  count(decade, flowering_range) %>%
+  ggplot(aes(x = decade, y = n, fill = flowering_range)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(
+    title = "Flowering Range by Decade",
+    x = "Decade",
+    y = "Count",
+    fill = "Flowering Range"
+  ) +
+  theme_minimal()
+
+plot_flowering_decade
+
 ```

### 2024-12-03 paper/paper.qmd
Results and Discussion draft addition

@@
 
-Using @fig-flowering-trend, we see a gradual shift in flowering dates over time.
-Earlier flowering events have become more common, particularly in modern years.
-This trend, captured by the loess regression, suggests an accelerating response to environmental changes.
-Similarly, @fig-temperature-trend reveals a steady increase in average March temperatures over the centuries, with notable warming in recent decades.
+@fig-flowering-trend shows the trend of sakura flowering dates across recorded history.
+The data presents consistency in flowering patterns until the late 20th century, with minor fluctuations attributable to natural climate variability.
+However, there was a shift in the modern era, where flowering dates have advanced significantly earlier in the year.
+This acceleration in phenological changes aligns with evidence of climate change, as warmer temperatures can progress earlier in flowering.
+The loess smoothing line emphasizes this trend, showing a stable period historically followed by a marked decline in flowering dates over the last century.
+
+@fig-temperature-trend captures the trajectory of average temperature over time, providing context for the shifts in flowering dates.
+Similar to the flowering date trends, temperature records remain relatively stable until the onset of industrialization.
+From approximately 1750 onward, a gradual warming trend ended in a sharp rise in the 20th century.
+This dramatic increase correlates with the earlier flowering patterns observed in @fig-flowering-trend.
 
@@
 
-## Relationship Between Temperature and Flowering Dates
+## Historical vs. Modern Data Comparison
 
-```{r}
-#| label: fig-temp-vs-flowering
-#| fig-cap: "Relationship Between Temperature and Flowering Date"
-#| warning: false 
-#| message: false
-#| fig-align: center
-
-plot_temp_vs_flowering <- ggplot(clean_sakura_data, aes(x = avg_temperature, y = day_of_year)) +
-  geom_point(alpha = 0.5, color = "purple") +
-  geom_smooth(method = "lm", color = "black", se = TRUE) +
-  labs(
-    title = "Relationship Between Temperature and Flowering Date",
-    x = "Average Temperature (°C)",
-    y = "Day of Year (Flowering Date)"
-  ) +
-  theme_minimal()
-
-plot_temp_vs_flowering
-
-```
-
-## Historical vs. Modern Data Comparison
+@fig-flowering-source compares flowering dates between historical and modern data sources.
+The boxplot reveals that historical flowering dates are more tightly clustered around the median, with less variability compared to modern data.
+The broader variability of flowering dates in modern data suggests that the sakura phenology is increasingly sensitive to environmental changes in the modern era.
 
@@
 
-## Flowering Ranges Across Decades
+# Discussion {#sec-discussion}
+The study revealed a consistent advancement in sakura flowering dates, particularly in modern data, driven by the rise in temperatures.
+Historical data showed stable patterns, while modern data indicated broader variability, likely due to several factors such as urbanization and environmental factors.
+These results highlight the sensitivity of sakura phenology to climate change and the increasing complexity of predicting the shift.
 
-```{r}
-#| label: fig-flowering-decade
-#| fig-cap: "Flowering Range by Decade"
-#| warning: false 
-#| message: false
-#| fig-align: center
+## Weaknesses
+Limitations include potential biases in historical data due to inconsistent and missing recording methods, geographic gaps, and modern data skewed toward urbanized areas.
+The analysis also lacked consideration of other environmental variables, such as precipitation and soil conditions, and other human interactions that could impact flowering dates.
 
-plot_flowering_decade <- clean_sakura_data %>%
-  count(decade, flowering_range) %>%
-  ggplot(aes(x = decade, y = n, fill = flowering_range)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(
-    title = "Flowering Range by Decade",
-    x = "Decade",
-    y = "Count",
-    fill = "Flowering Range"
-  ) +
-  theme_minimal()
-
-plot_flowering_decade
-
-```
-
-# Discussion {#sec-discussion}
+## Future Directions 
+Future research could expand on considering human factors and other geographical factors.
+Incorporating variables like precipitation, human movement in the area, and integrating climate models could improve predictive accuracy.
 

### 2024-12-03 other/datasheet/datasheet.qmd
add datasheet

---
title: "Datasheet for 'Analyzing Sakura Flowering Dates and Climate Trends'"
subtitle: "A Phenological Study Using Historical and Modern Data"
author: 
  - Mariko Lee
thanks: "Code and data are available at: [https://github.com/leemarik/sakura-flowering-trends.git]."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "This datasheet documents a dataset on sakura flowering dates and associated climate variables. The dataset integrates historical and modern records to examine phenological shifts driven by climate change, analyzing the impact of rising temperatures on sakura flowering patterns."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of sakura flowering trends. It addresses the gap in structured, longitudinal data combining historical and modern phenological records.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The `sakura-historical` dataset is from Prof. Yasuyuki Aono from Osaka Metropolitan University and `sakura-modern` and `temperature-modern` dataset is from the Japan meteorological Agency. 
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The dataset relies on publicly available historical records and institutional data, requiring no external funding.
4. *Any other comments?*
    - The dataset are retrieved from Alex Cookson [https://github.com/tacookson/data/tree/master/sakura-flowering] GitHub.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- The primary target variable is `day_of_year`.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some March temperature data is unavailable for historical observations due to gaps in records.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- The dataset links flowering dates to climatic variables and geographic regions.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- The dataset is split into 80% training and 20% testing subsets for modeling purposes.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Historical data may include biases from estimation methods. Modern data may reflect urban heat island effects.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Data sources include: Prof. Yasuyuki Aono’s Osaka Metropolitan University　生態気象学研究グループ website (@yasuyukiphenology15), Japan Meteorological Agency (JMA) phenological and climate statistics (@JapanMeteorAgency), NOAA’s National Centers for Environmental Information (@ncei2024)
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-12-03 other/datasheet/datasheet.qmd
dataset reference fix

@@
 	- Data sources include: Prof.
-Yasuyuki Aono’s Osaka Metropolitan University　生態気象学研究グループ website (@yasuyukiphenology15), Japan Meteorological Agency (JMA) phenological and climate statistics (@JapanMeteorAgency), NOAA’s National Centers for Environmental Information (@ncei2024)
+Yasuyuki Aono’s website (@yasuyukiphenology15), Japan Meteorological Agency (JMA) phenological and climate statistics (@JapanMeteorAgency), NOAA’s National Centers for Environmental Information (@ncei2024)
 11.

### 2024-12-03 other/datasheet/datasheet.qmd
Completed Datasheet update

@@
 thanks: "Code and data are available at: [https://github.com/leemarik/sakura-flowering-trends.git]."
-date: "`r Sys.time()`"
+date: today
 date-format: "D MMMM YYYY"
@@
 Please provide a description.*
-	- TBD
+	- Each instance represents a sakura flowering observation, along with corresponding temperature data and metadata (e.g., year, source).
 2. *How many instances are there in total (of each type, if appropriate)?*
-	- TBD
+	- The dataset contains 3,526 observations spanning over 1,200 years.
 3.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-	- TBD
+	- The dataset includes historical records for Kyoto and modern observations across Japan. While not exhaustive, it provides broad geographic and temporal coverage.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
-	- TBD
+	- Instances consist of variables such as flowering date (doy), full bloom date, monthly mean temp, and geographic coordinates (latitude, longitude)
 5. *Is there a label or target associated with each instance? If so, please provide a description.*
-	- The primary target variable is `day_of_year`.
+	- The primary target variable is flowering date `day_of_year`.
 6.
@@
 If so, please provide a description.*
-	- TBD
+	- No, the dataset is public
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- TBD
+	- The dataset focuses on ecological phenomena and does not contain sensitive personal data.
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+	- Historical data is limited to Kyoto, while modern data covers a broader geographic range.
 14.
@@
 If so, please describe how.*
-	- TBD
+	- No personal information is included in the dataset.
 15.
@@
 If so, please provide a description.*
-	- TBD
+	- The dataset does not raise ethical concerns or pose risks of harm or offense.
 16. *Any other comments?*
-	- TBD
+	- No
 
@@
 If so, please describe how.*
-	- TBD
+	- Historical flowering data was extracted from Prof. Yasuyuki Aono’s studies and modern data from JMA records, collected by Alex Cookson. 
 2.
@@
 How were these mechanisms or procedures validated?*
-	- TBD
+	- Data was collected manually and through automated extraction methods.
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+	- Comprehensive for Kyoto historical data and geographically representative for modern Japan.
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+	- Historical data was compiled by researchers; modern data was sourced from JMA records.
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+	- The historical dataset spans 812–2015, and the modern dataset spans 1953–2019.
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- No review was necessary as the dataset does not involve human subjects or private information.
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+	- No, data was obtained from publicly available research and meteorological records.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+	- N/A; the dataset does not involve human participants.
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+	- N/A
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+	- N/A
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- The dataset supports climate research without risks to individuals or groups.
 12. *Any other comments?*
-	- TBD
+	- No
 
@@
 If not, you may skip the remaining questions in this section.*
-	- TBD
+	- Missing values were filtered, and flowering dates were standardized into `day_of_year` format for consistency. Historical estimates were aligned with modern records where possible.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+	- The cleaned dataset is available in `.parquet` and `.csv` formats. It is avialable at [https://github.com/leemarik/sakura-flowering-trends/tree/master/data/01-raw_data]
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
+	- R was used for preprocessing, employing packages like `tidyverse`, `lubridate`, and `arrow`.
+Data Cleaning script is available at [https://github.com/leemarik/sakura-flowering-trends/blob/master/scripts/03-clean_data.R]
 4. *Any other comments?*
-	- TBD
+	- No
 
@@
 1. *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- TBD
+	- The dataset has been used to model trends in sakura flowering dates using linear regression and generalized additive models (GAMs).
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- TBD
+	- Yes, the dataset and analysis code are hosted on GitHub at [https://github.com/leemarik/sakura-flowering-trends]
 3. *What (other) tasks could the dataset be used for?*
-	- TBD
+	- It can be used for climate forecasting, ecological impact analysis, and cultural studies.
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
+	- Dataset consumers should account for biases in historical data and the limitations of temperature proxies.
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- No
 6. *Any other comments?*
-	- TBD
+	- No
 
@@
 If so, please provide a description.*
-	- TBD
+	- The dataset is publicly available at [https://github.com/leemarik/sakura-flowering-trends.git]
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+	- Available via GitHub as `.csv` and `.parquet` files with documentation.
 3. *When will the dataset be distributed?*
-	- TBD
+	- Distributed upon project completion.
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+	- No. MIT License
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+	- No
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
+	- No
 7. *Any other comments?*
-	- TBD
+	- No
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+	- The dataset will be maintained by Mariko Lee during the project term.
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+	- For inquiries, contact mariko.lee@mail.utoronto.ca
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+	- No
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+	- No
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+	- N/A
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+	- No, as the dataset is updated. However, older version will be accessible through update history through GitHub. 
 7.
@@
 If so, please provide a description.*
-	- TBD
+	- Contributions are welcome via GitHub pull requests.
 8. *Any other comments?*
-	- TBD
+	- No
 

### 2024-12-03 scripts/00-simulate_data.R
Add sketches

@@ -1,70 +1,0 @@
-#### Preamble ####
-# Purpose: Simulate the analysis dataset
-# Author: Mariko Lee
-# Date: 19 November 2024
-# Contact: mariko.lee@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-
-#### Workspace setup ####
-library(tidyverse)
-library(lubridate)
-library(arrow)
-
-#### Parameters for Simulation ####
-set.seed(123)  
-
-# Define the number of years for simulation
-n_years <- 1000 
-start_year <- 800
-end_year <- start_year + n_years - 1
-
-# Define temperature range and relationship
-avg_temp_range <- c(5, 15)  
-temp_effect_on_flowering <- -3  
-baseline_flowering_day <- 100  
-
-#### Simulate Data ####
-
-# Simulate year
-year <- seq(start_year, end_year)
-
-# Simulate average temperature
-avg_temperature <- runif(n_years, min = avg_temp_range[1], max = avg_temp_range[2])
-
-# Simulate flowering day of year
-day_of_year <- baseline_flowering_day + temp_effect_on_flowering * (avg_temperature - mean(avg_temp_range))
-day_of_year <- round(day_of_year) 
-
-# Simulate decade
-decade <- floor(year / 10) * 10
-
-# Simulate flowering range categories
-flowering_range <- case_when(
-  day_of_year <= 90 ~ "Early (Jan-Mar)",
-  day_of_year > 90 & day_of_year <= 120 ~ "Mid (Apr)",
-  day_of_year > 120 ~ "Late (May+)"
-)
-
-# Simulate lat/lon (random location in Kyoto area for demonstration)
-lat <- runif(n_years, min = 34.9, max = 35.0)  # Latitude range
-lon <- runif(n_years, min = 135.7, max = 135.8)  # Longitude range
-
-# Source type (randomly assign historical/modern)
-source <- sample(c("historical", "modern"), n_years, replace = TRUE, prob = c(0.7, 0.3))
-
-#### Combine Simulated Data ####
-simulated_data <- tibble(
-  year = year,
-  avg_temperature = avg_temperature,
-  day_of_year = day_of_year,
-  flowering_range = flowering_range,
-  lat = lat,
-  lon = lon,
-  decade = decade,
-  source = source
-)
-
-
-#### Save data ####
-write_csv(simulated_data, "data/00-simulated_data/simulated_sakura_data.csv")

### 2024-12-03 scripts/01-test_simulated_data.R

@@ -1,74 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the structure and validity of the Simulated Data.
-# Author: Mariko Lee
-# Date: 19 November 2024
-# Contact: mariko.lee@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? None
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-
-#### Load simulated data ####
-simulated_data <- read_csv("data/00-simulated_data/simulated_sakura_data.csv")
-
-#### Test if the data was successfully loaded ####
-if (exists("simulated_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
-
-#### Test Simulated Data ####
-
-# Expected ranges and categories
-expected_flowering_ranges <- c("Early (Jan-Mar)", "Mid (Apr)", "Late (May+)")
-expected_sources <- c("historical", "modern")
-expected_lat_range <- c(34.9, 35.0)  # Kyoto latitude range
-expected_lon_range <- c(135.7, 135.8)  # Kyoto longitude range
-expected_temp_range <- c(5, 15)  # Temperature range
-expected_day_of_year_range <- c(1, 365)  # Valid day of year range
-
-# Test 1: Check for missing values
-cat("Checking for missing values...\n")
-missing_values <- simulated_data %>%
-  summarise(across(everything(), ~ sum(is.na(.)))) %>%
-  pivot_longer(cols = everything(), names_to = "variable", values_to = "missing_count")
-print(missing_values)
-stopifnot(all(missing_values$missing_count == 0))
-
-# Test 2: Validate `avg_temperature` range
-cat("Validating avg_temperature range...\n")
-stopifnot(all(simulated_data$avg_temperature >= expected_temp_range[1] & 
-                simulated_data$avg_temperature <= expected_temp_range[2]))
-
-# Test 3: Validate `day_of_year` range
-cat("Validating day_of_year range...\n")
-stopifnot(all(simulated_data$day_of_year >= expected_day_of_year_range[1] & 
-                simulated_data$day_of_year <= expected_day_of_year_range[2]))
-
-# Test 4: Validate `flowering_range` categories
-cat("Validating flowering_range categories...\n")
[…]

### 2024-12-03 scripts/03-clean_data.R

@@ -1,82 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw data 
-# Author: Mariko Lee
-# Date: 19 November 2024
-# Contact: mariko.lee@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-
-#### Workspace setup ####
-library(tidyverse)
-library(lubridate)
-library(arrow)
-
-# Load datasets
-historical <- read_csv("data/01-raw_data/sakura-historical.csv")
-modern <- read_csv("data/01-raw_data/sakura-modern.csv")
-temperatures <- read_csv("data/01-raw_data/temperatures-modern.csv")
-
-# Clean historical dataset
-historical <- historical %>%
-  select(year, flower_date, temp_c_recon) %>%
-  rename(
-    flowering_date = flower_date,
-    temperature_recon = temp_c_recon
-  ) %>%
-  mutate(
-    flowering_date = ymd(flowering_date),
-    source = "historical"
-  )
-
-# Clean modern dataset
-modern <- modern %>%
-  select(year, flower_date, full_bloom_date) %>%  # Exclude lat and lon
-  rename(flowering_date = flower_date) %>%
-  mutate(
-    flowering_date = ymd(flowering_date),
-    source = "modern"
-  )
-
-# Clean temperatures dataset
-temperatures <- temperatures %>%
-  filter(month == "Mar") %>%
-  group_by(year) %>%
-  summarize(avg_temperature = mean(mean_temp_c, na.rm = TRUE), .groups = "drop")
-
-# Combine historical and modern flowering data
-clean_sakura_data <- bind_rows(historical, modern)
-
-# Join with temperature data
-clean_sakura_data <- clean_sakura_data %>%
-  left_join(temperatures, by = "year")
-
-# Fill missing avg_temperature with temperature_recon
-clean_sakura_data <- clean_sakura_data %>%
-  mutate(avg_temperature = ifelse(is.na(avg_temperature), temperature_recon, avg_temperature))
-
-# Remove rows with missing flowering dates or avg_temperatures
-clean_sakura_data <- clean_sakura_data %>%
-  filter(!is.na(flowering_date) & !is.na(avg_temperature))
-
-# Add derived columns
-clean_sakura_data <- clean_sakura_data %>%
-  mutate(
-    day_of_year = yday(flowering_date),
-    decade = floor(year / 10) * 10,
-    flowering_range = case_when(
[…]

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -1,101 +1,0 @@
-#### Preamble ####
-# Purpose: Tests for Cleaned Dataset
-# Author: Mariko Lee
-# Date: 19 November 2024
-# Contact: mariko.lee@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Run 03-clean_data.R script
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-
-#### Test Data ####
-clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
-
-#### Initialize validation flag ####
-all_tests_passed <- TRUE
-
-# Test 1: Check for missing values
-cat("Test 1: Checking for missing values...\n")
-missing_values <- clean_sakura_data %>%
-  summarise(across(everything(), ~ sum(is.na(.)))) %>%
-  pivot_longer(cols = everything(), names_to = "variable", values_to = "missing_count")
-print(missing_values)
-if (any(missing_values$missing_count > 0)) {
-  cat("FAILED: Missing values found in the dataset.\n")
-  all_tests_passed <- FALSE
-} else {
-  cat("PASSED: No missing values found.\n")
-}
-
-# Test 2: Validate column count
-cat("\nTest 2: Validating column count...\n")
-expected_columns <- c("year", "flowering_date", "avg_temperature",
-                      "day_of_year", "decade", "flowering_range", "source")
-if (ncol(clean_sakura_data) != length(expected_columns)) {
-  cat("FAILED: Column count mismatch. Expected:", length(expected_columns), "but found:", ncol(clean_sakura_data), "\n")
-  all_tests_passed <- FALSE
-} else {
-  cat("PASSED: Column count matches expectations.\n")
-}
-
-# Test 3: Validate `avg_temperature` range
-cat("\nTest 3: Validating avg_temperature range...\n")
-if (!all(clean_sakura_data$avg_temperature >= 5 & clean_sakura_data$avg_temperature <= 15)) {
-  cat("FAILED: Some avg_temperature values are outside the valid range (5-15°C).\n")
-  all_tests_passed <- FALSE
-} else {
-  cat("PASSED: All avg_temperature values are within the valid range.\n")
-}
-
-# Test 4: Validate `day_of_year` range
-cat("\nTest 4: Validating day_of_year range...\n")
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -1,129 +1,0 @@
-#### Preamble ####
-# Purpose: Models 
-# Author: Mariko Lee
-# Date: 19 November 2024
-# Contact: mariko.lee@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-
-
-#### Workspace setup ####
-library(tidyverse)  
-library(arrow)
-library(patchwork)
-library(ggplot2)
-
-#### Read data ####
-clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
-
-#### Summary Statistics ####
-# Summary of the key variables
-summary_stats <- clean_sakura_data %>%
-  summarize(
-    min_year = min(year),
-    max_year = max(year),
-    avg_temp_mean = mean(avg_temperature, na.rm = TRUE),
-    avg_temp_sd = sd(avg_temperature, na.rm = TRUE),
-    day_of_year_mean = mean(day_of_year, na.rm = TRUE),
-    day_of_year_sd = sd(day_of_year, na.rm = TRUE)
-  )
-print(summary_stats)
-
-#### Visualizations ####
-
-# 1. Distribution of Flowering Dates (day_of_year)
-plot_flowering_distribution <- ggplot(clean_sakura_data, aes(x = day_of_year)) +
-  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
-  labs(
-    title = "Distribution of Sakura Flowering Dates",
-    x = "Day of Year (Flowering Date)",
-    y = "Frequency"
-  ) +
-  theme_minimal()
-
-# 2. Average Temperature Distribution
-plot_temperature_distribution <- ggplot(clean_sakura_data, aes(x = avg_temperature)) +
-  geom_histogram(binwidth = 1, fill = "coral", color = "black") +
-  labs(
-    title = "Distribution of Average Temperatures",
-    x = "Average Temperature (°C)",
-    y = "Frequency"
-  ) +
-  theme_minimal()
-
-# 3. Trend of Flowering Dates Over Time
-plot_flowering_trend <- ggplot(clean_sakura_data, aes(x = year, y = day_of_year)) +
-  geom_point(alpha = 0.5, color = "darkgreen") +
-  geom_smooth(method = "loess", color = "blue", se = TRUE) +
-  labs(
-    title = "Trend of Sakura Flowering Dates Over Time",
-    x = "Year",
-    y = "Day of Year (Flowering Date)"
-  ) +
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -1,94 +1,0 @@
-#### Preamble ####
-# Purpose: Model
-# Author: Mariko Lee
-# Date: 19 November 2024
-# Contact: mariko.lee@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-
-#### Workspace setup ####
-library(tidyverse)
-library(mgcv)  # For GAM modeling
-library(arrow) # For data input/output
-library(broom) # For tidying model results
-
-#### Read cleaned data ####
-clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
-
-#### Prepare data ####
-# Filter data to ensure no missing critical variables
-filtered_data <- clean_sakura_data %>%
-  filter(!is.na(avg_temperature) & !is.na(day_of_year))
-
-#### Baseline Model: Linear Regression ####
-# Linear regression with temperature and year as predictors
-lm_model <- lm(day_of_year ~ avg_temperature + year, data = filtered_data)
-
-# Summary of the Linear Regression model
-cat("\n--- Linear Regression Model Summary ---\n")
-summary(lm_model)
-
-# Diagnostics: Residual Plot
-par(mfrow = c(2, 2))
-plot(lm_model)
-
-# Tidy the linear regression model output for interpretation
-lm_results <- tidy(lm_model)
-cat("\n--- Linear Regression Coefficients ---\n")
-print(lm_results)
-
-#### Generalized Additive Model (GAM) ####
-# GAM model with smooth terms for temperature and year
-gam_model <- gam(day_of_year ~ s(avg_temperature) + s(year),
-                 data = filtered_data,
-                 method = "REML")  # Restricted Maximum Likelihood
-
-# Summary of the GAM model
-cat("\n--- GAM Model Summary ---\n")
-summary(gam_model)
-
-# Diagnostics: GAM Residuals and Smooth Terms
-par(mfrow = c(2, 2))
-gam.check(gam_model)
-
-# Plot the smooth terms for temperature and year
-par(mfrow = c(1, 2))
-plot(gam_model, select = 1, shade = TRUE, rug = TRUE, seWithMean = TRUE, main = "Effect of Temperature")
-plot(gam_model, select = 2, shade = TRUE, rug = TRUE, seWithMean = TRUE, main = "Effect of Year")
-
-#### Compare Models ####
-# Compare models based on AIC (Akaike Information Criterion)
-cat("\n--- Model Comparison ---\n")
[…]

### 2024-12-03 scripts/07-train_data.R

@@ -1,30 +1,0 @@
-#### Preamble ####
-# Purpose: testing and training data
-# Author: Mariko Lee
-# Date: 19 November 2024
-# Contact: mariko.lee@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-
-
-#### Workspace setup ####
-library(dplyr)
-library(rsample)
-library(arrow)
-
-# Read the cleaned sakura dataset
-clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
-
-# Set seed for reproducibility
-set.seed(123)  
-
-# Split the data into 80% training and 20% testing
-split <- initial_split(clean_sakura_data, prop = 0.8)
-
-# Extract training and testing datasets
-train_data <- training(split)
-test_data <- testing(split)
-
-# Save training and testing datasets as Parquet files
-write_parquet(train_data, "data/03-model_data/train_data.parquet")
-write_parquet(test_data, "data/03-model_data/test_data.parquet")

### 2024-12-03 paper/paper.qmd
add

@@
 #| fig-cap: "Distribution of Sakura Flowering Dates"
+#| echo: false
 #| warning: false 
@@
 #| fig-cap: "Distribution of Average Temperatures in March"
+#| echo: false
 #| warning: false 
@@
 #| fig-cap: "Trend of Sakura Flowering Dates Over Time"
+#| echo: false
 #| warning: false 
@@
 #| fig-cap: "Trend of Average Temperatures Over Time"
+#| echo: false
 #| warning: false 
@@
 #| fig-align: center
-#| 
+
 plot_temperature_trend <- ggplot(clean_sakura_data, aes(x = year, y = avg_temperature)) +
@@
 #| fig-cap: "Comparison of Flowering Dates by Data Source"
+#| echo: false
 #| warning: false 

### 2024-12-03 scripts/00-simulate_data.R

@@ -1,0 +1,70 @@
+#### Preamble ####
+# Purpose: Simulate the analysis dataset
+# Author: Mariko Lee
+# Date: 19 November 2024
+# Contact: mariko.lee@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+library(tidyverse)
+library(lubridate)
+library(arrow)
+
+#### Parameters for Simulation ####
+set.seed(123)  
+
+# Define the number of years for simulation
+n_years <- 1000 
+start_year <- 800
+end_year <- start_year + n_years - 1
+
+# Define temperature range and relationship
+avg_temp_range <- c(5, 15)  
+temp_effect_on_flowering <- -3  
+baseline_flowering_day <- 100  
+
+#### Simulate Data ####
+
+# Simulate year
+year <- seq(start_year, end_year)
+
+# Simulate average temperature
+avg_temperature <- runif(n_years, min = avg_temp_range[1], max = avg_temp_range[2])
+
+# Simulate flowering day of year
+day_of_year <- baseline_flowering_day + temp_effect_on_flowering * (avg_temperature - mean(avg_temp_range))
+day_of_year <- round(day_of_year) 
+
+# Simulate decade
+decade <- floor(year / 10) * 10
+
+# Simulate flowering range categories
+flowering_range <- case_when(
+  day_of_year <= 90 ~ "Early (Jan-Mar)",
+  day_of_year > 90 & day_of_year <= 120 ~ "Mid (Apr)",
+  day_of_year > 120 ~ "Late (May+)"
+)
+
+# Simulate lat/lon (random location in Kyoto area for demonstration)
+lat <- runif(n_years, min = 34.9, max = 35.0)  # Latitude range
+lon <- runif(n_years, min = 135.7, max = 135.8)  # Longitude range
+
+# Source type (randomly assign historical/modern)
+source <- sample(c("historical", "modern"), n_years, replace = TRUE, prob = c(0.7, 0.3))
+
+#### Combine Simulated Data ####
+simulated_data <- tibble(
+  year = year,
+  avg_temperature = avg_temperature,
+  day_of_year = day_of_year,
+  flowering_range = flowering_range,
+  lat = lat,
+  lon = lon,
+  decade = decade,
+  source = source
+)
+
+
+#### Save data ####
+write_csv(simulated_data, "data/00-simulated_data/simulated_sakura_data.csv")

### 2024-12-03 scripts/01-test_simulated_data.R

@@ -1,0 +1,74 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the Simulated Data.
+# Author: Mariko Lee
+# Date: 19 November 2024
+# Contact: mariko.lee@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - 00-simulate_data.R must have been run
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+
+#### Load simulated data ####
+simulated_data <- read_csv("data/00-simulated_data/simulated_sakura_data.csv")
+
+#### Test if the data was successfully loaded ####
+if (exists("simulated_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+#### Test Simulated Data ####
+
+# Expected ranges and categories
+expected_flowering_ranges <- c("Early (Jan-Mar)", "Mid (Apr)", "Late (May+)")
+expected_sources <- c("historical", "modern")
+expected_lat_range <- c(34.9, 35.0)  # Kyoto latitude range
+expected_lon_range <- c(135.7, 135.8)  # Kyoto longitude range
+expected_temp_range <- c(5, 15)  # Temperature range
+expected_day_of_year_range <- c(1, 365)  # Valid day of year range
+
+# Test 1: Check for missing values
+cat("Checking for missing values...\n")
+missing_values <- simulated_data %>%
+  summarise(across(everything(), ~ sum(is.na(.)))) %>%
+  pivot_longer(cols = everything(), names_to = "variable", values_to = "missing_count")
+print(missing_values)
+stopifnot(all(missing_values$missing_count == 0))
+
+# Test 2: Validate `avg_temperature` range
+cat("Validating avg_temperature range...\n")
+stopifnot(all(simulated_data$avg_temperature >= expected_temp_range[1] & 
+                simulated_data$avg_temperature <= expected_temp_range[2]))
+
+# Test 3: Validate `day_of_year` range
+cat("Validating day_of_year range...\n")
+stopifnot(all(simulated_data$day_of_year >= expected_day_of_year_range[1] & 
+                simulated_data$day_of_year <= expected_day_of_year_range[2]))
+
+# Test 4: Validate `flowering_range` categories
+cat("Validating flowering_range categories...\n")
[…]

### 2024-12-03 scripts/03-clean_data.R

@@ -1,0 +1,82 @@
+#### Preamble ####
+# Purpose: Cleans the raw data 
+# Author: Mariko Lee
+# Date: 19 November 2024
+# Contact: mariko.lee@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+library(tidyverse)
+library(lubridate)
+library(arrow)
+
+# Load datasets
+historical <- read_csv("data/01-raw_data/sakura-historical.csv")
+modern <- read_csv("data/01-raw_data/sakura-modern.csv")
+temperatures <- read_csv("data/01-raw_data/temperatures-modern.csv")
+
+# Clean historical dataset
+historical <- historical %>%
+  select(year, flower_date, temp_c_recon) %>%
+  rename(
+    flowering_date = flower_date,
+    temperature_recon = temp_c_recon
+  ) %>%
+  mutate(
+    flowering_date = ymd(flowering_date),
+    source = "historical"
+  )
+
+# Clean modern dataset
+modern <- modern %>%
+  select(year, flower_date, full_bloom_date) %>%  # Exclude lat and lon
+  rename(flowering_date = flower_date) %>%
+  mutate(
+    flowering_date = ymd(flowering_date),
+    source = "modern"
+  )
+
+# Clean temperatures dataset
+temperatures <- temperatures %>%
+  filter(month == "Mar") %>%
+  group_by(year) %>%
+  summarize(avg_temperature = mean(mean_temp_c, na.rm = TRUE), .groups = "drop")
+
+# Combine historical and modern flowering data
+clean_sakura_data <- bind_rows(historical, modern)
+
+# Join with temperature data
+clean_sakura_data <- clean_sakura_data %>%
+  left_join(temperatures, by = "year")
+
+# Fill missing avg_temperature with temperature_recon
+clean_sakura_data <- clean_sakura_data %>%
+  mutate(avg_temperature = ifelse(is.na(avg_temperature), temperature_recon, avg_temperature))
+
+# Remove rows with missing flowering dates or avg_temperatures
+clean_sakura_data <- clean_sakura_data %>%
+  filter(!is.na(flowering_date) & !is.na(avg_temperature))
+
+# Add derived columns
+clean_sakura_data <- clean_sakura_data %>%
+  mutate(
+    day_of_year = yday(flowering_date),
+    decade = floor(year / 10) * 10,
+    flowering_range = case_when(
[…]

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -1,0 +1,101 @@
+#### Preamble ####
+# Purpose: Tests for Cleaned Dataset
+# Author: Mariko Lee
+# Date: 19 November 2024
+# Contact: mariko.lee@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Run 03-clean_data.R script
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+
+#### Test Data ####
+clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
+
+#### Initialize validation flag ####
+all_tests_passed <- TRUE
+
+# Test 1: Check for missing values
+cat("Test 1: Checking for missing values...\n")
+missing_values <- clean_sakura_data %>%
+  summarise(across(everything(), ~ sum(is.na(.)))) %>%
+  pivot_longer(cols = everything(), names_to = "variable", values_to = "missing_count")
+print(missing_values)
+if (any(missing_values$missing_count > 0)) {
+  cat("FAILED: Missing values found in the dataset.\n")
+  all_tests_passed <- FALSE
+} else {
+  cat("PASSED: No missing values found.\n")
+}
+
+# Test 2: Validate column count
+cat("\nTest 2: Validating column count...\n")
+expected_columns <- c("year", "flowering_date", "avg_temperature",
+                      "day_of_year", "decade", "flowering_range", "source")
+if (ncol(clean_sakura_data) != length(expected_columns)) {
+  cat("FAILED: Column count mismatch. Expected:", length(expected_columns), "but found:", ncol(clean_sakura_data), "\n")
+  all_tests_passed <- FALSE
+} else {
+  cat("PASSED: Column count matches expectations.\n")
+}
+
+# Test 3: Validate `avg_temperature` range
+cat("\nTest 3: Validating avg_temperature range...\n")
+if (!all(clean_sakura_data$avg_temperature >= 5 & clean_sakura_data$avg_temperature <= 15)) {
+  cat("FAILED: Some avg_temperature values are outside the valid range (5-15°C).\n")
+  all_tests_passed <- FALSE
+} else {
+  cat("PASSED: All avg_temperature values are within the valid range.\n")
+}
+
+# Test 4: Validate `day_of_year` range
+cat("\nTest 4: Validating day_of_year range...\n")
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -1,0 +1,129 @@
+#### Preamble ####
+# Purpose: Models 
+# Author: Mariko Lee
+# Date: 19 November 2024
+# Contact: mariko.lee@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+
+#### Workspace setup ####
+library(tidyverse)  
+library(arrow)
+library(patchwork)
+library(ggplot2)
+
+#### Read data ####
+clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
+
+#### Summary Statistics ####
+# Summary of the key variables
+summary_stats <- clean_sakura_data %>%
+  summarize(
+    min_year = min(year),
+    max_year = max(year),
+    avg_temp_mean = mean(avg_temperature, na.rm = TRUE),
+    avg_temp_sd = sd(avg_temperature, na.rm = TRUE),
+    day_of_year_mean = mean(day_of_year, na.rm = TRUE),
+    day_of_year_sd = sd(day_of_year, na.rm = TRUE)
+  )
+print(summary_stats)
+
+#### Visualizations ####
+
+# 1. Distribution of Flowering Dates (day_of_year)
+plot_flowering_distribution <- ggplot(clean_sakura_data, aes(x = day_of_year)) +
+  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
+  labs(
+    title = "Distribution of Sakura Flowering Dates",
+    x = "Day of Year (Flowering Date)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+# 2. Average Temperature Distribution
+plot_temperature_distribution <- ggplot(clean_sakura_data, aes(x = avg_temperature)) +
+  geom_histogram(binwidth = 1, fill = "coral", color = "black") +
+  labs(
+    title = "Distribution of Average Temperatures",
+    x = "Average Temperature (°C)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+# 3. Trend of Flowering Dates Over Time
+plot_flowering_trend <- ggplot(clean_sakura_data, aes(x = year, y = day_of_year)) +
+  geom_point(alpha = 0.5, color = "darkgreen") +
+  geom_smooth(method = "loess", color = "blue", se = TRUE) +
+  labs(
+    title = "Trend of Sakura Flowering Dates Over Time",
+    x = "Year",
+    y = "Day of Year (Flowering Date)"
+  ) +
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -1,0 +1,94 @@
+#### Preamble ####
+# Purpose: Model
+# Author: Mariko Lee
+# Date: 19 November 2024
+# Contact: mariko.lee@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+library(tidyverse)
+library(mgcv)  # For GAM modeling
+library(arrow) # For data input/output
+library(broom) # For tidying model results
+
+#### Read cleaned data ####
+clean_sakura_data <- read_parquet("data/02-analysis_data/clean_sakura_data.parquet")
+
+#### Prepare data ####
+# Filter data to ensure no missing critical variables
+filtered_data <- clean_sakura_data %>%
+  filter(!is.na(avg_temperature) & !is.na(day_of_year))
+
+#### Baseline Model: Linear Regression ####
+# Linear regression with temperature and year as predictors
+lm_model <- lm(day_of_year ~ avg_temperature + year, data = filtered_data)
+
+# Summary of the Linear Regression model
+cat("\n--- Linear Regression Model Summary ---\n")
+summary(lm_model)
+
+# Diagnostics: Residual Plot
+par(mfrow = c(2, 2))
+plot(lm_model)
+
+# Tidy the linear regression model output for interpretation
+lm_results <- tidy(lm_model)
+cat("\n--- Linear Regression Coefficients ---\n")
+print(lm_results)
+
+#### Generalized Additive Model (GAM) ####
+# GAM model with smooth terms for temperature and year
+gam_model <- gam(day_of_year ~ s(avg_temperature) + s(year),
+                 data = filtered_data,
+                 method = "REML")  # Restricted Maximum Likelihood
+
+# Summary of the GAM model
+cat("\n--- GAM Model Summary ---\n")
+summary(gam_model)
+
+# Diagnostics: GAM Residuals and Smooth Terms
+par(mfrow = c(2, 2))
+gam.check(gam_model)
+
+# Plot the smooth terms for temperature and year
+par(mfrow = c(1, 2))
+plot(gam_model, select = 1, shade = TRUE, rug = TRUE, seWithMean = TRUE, main = "Effect of Temperature")
+plot(gam_model, select = 2, shade = TRUE, rug = TRUE, seWithMean = TRUE, main = "Effect of Year")
+
+#### Compare Models ####
+# Compare models based on AIC (Akaike Information Criterion)
+cat("\n--- Model Comparison ---\n")
[…]

### 2024-12-04 paper/paper.qmd
Final paper update

@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 @sec-results presents the results, followed by a discussion in @sec-discussion on the implications, limitations, and future directions.
+Further details on model and dataset will be discussed in @sec-appendix.
 
@@
 **Model Performance Metrics:** 
-For the linear regression model, the adjusted `R^2` was 0.0136, indicating a weak fit to the data.
+For the linear regression model, the adjusted $R^{2}$ was 0.0136, indicating a weak fit to the data.
 The residual standard error was 28.18 days, highlighting high variability in flowering dates not explained by the model.
-The GAM improved the adjusted `R^2` to 0.0411, explaining 4.36% of the deviance.
+The GAM improved the adjusted $R^{2}$ to 0.0411, explaining 4.36% of the deviance.
 While small, this improvement shows the GAM’s ability to capture non-linear trends lacking in the linear model.
@@
 Future predictions were based on three hypothetical warming scenarios:
+
 - +1°C: A moderate temperature increase.
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
+
+## Survey, Sampling, and Observational Data
+The analysis of the Sakura flowering trend explores phenological changes in sakura flowering dates using a combination of historical and modern datasets.
+Below, I detail the sampling methodologies and observational techniques.
+
+### Historical Data
+The historical sakura flowering dataset relies on secondary observations documented in diaries, hanami (flower-viewing) records, and cultural archives from Kyoto, dating back to 812.
+The flowering dates were reconstructed by paleoclimatologist Prof.
+Yasuyuki Aono (@yasuyukiphenology15), who used these observations to estimate March temperatures.
+These records represent non-probabilistic sampling, as they reflect the specific cultural practices of elite Japanese society and their interaction with nature.
+The dataset’s strengths include its temporal depth of the unique 1,200-year timeline and the cultural significance of its direct connections to human activity, making it highly relevant for climate impact studies.
+However, limitations arise from geographic bias, where the data is exclusively from Kyoto, limiting generalizability and sampling gaps, where variations in recordkeeping over centuries, including periods with sparse documentation.
+
+### Modern Data
+Modern flowering records from the Japan Meteorological Agency (@JapanMeteorAgency) cover nationwide observations from 1953 to 2019.
+JMA uses systematic, probabilistic sampling by selecting stations across Japan to record flowering dates and related phenological phenomena.
+Key aspects include consistency of standardized protocols for determining the “first flowering” stage and the broad coverage where data spans multiple climatic zones, from Kyushu in the south to Hokkaido in the north.
+
+### Observational Data Methods
+Modern observational data incorporates measurements of Average Monthly Temperatures collected from the Japan Meteorological Agency’s Station Data (@JapanMeteorAgency).
+Phenological Markers were defined by the ‘Biological Phenological Observation Guidelines’ to ensure consistency in reporting flowering stages.
+
+To validate the reliability of these observational methods, posterior predictive checks were applied to ensure consistency in modern data trends.
+
+## Simulations
+Simulations of future flowering dates (+1°C, +2°C, +3°C scenarios) rely on existing literature linking phenology to climate (@yasuyukiphenology15).
+These simulations emphasize the sensitivity of flowering dates to March temperatures, corroborating findings from prior studies.
+The results align with paleoclimatic models used in historical reconstructions, bridging the gap between observational and modeled data.
+
+#### Predictive Simulations
+Under hypothetical temperature increases, predictions showed a clear advancement in flowering dates:
+
+- +1°C: Flowering dates shifted earlier by 5–7 days.
+- +2°C: Shifts of 10–12 days were observed.
+- +3°C: The most dramatic change, with flowering advancing by up to 15 days.
+
+These simulations validate the model's ability to capture the non-linear relationships between temperature and phenological changes.
 
